Compositional equivalences based on Open pNets

### 2020-04-28 2016-Weak-FH-Bisim/WeakBisim.tex
variable simplification everywhere

We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:
\begin{lemma}[Open transition decomposition\label{lem-decompose}] Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

More precisely, we state that $\tau$ is not observable if the automaton always allows any $\tau$ transition from holes, and additionally the global transition resulting from a $\tau$ action of a hole is a $\tau$ transition not changing the pNet's state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of $\tau$ actions]\label{def:Non-ObsTau}
An open automaton $A = <J,\mathcal{S},s_0,V,\mathcal{T}>$ \emph{cannot observe $\tau$ actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have:
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\beta_j$, $J$,  $\alpha$,  $s$, $s'$, $\Pred$, $\Post$  such that
\[ \openrule
         {
           \beta_j^{j\in J},\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] If there exists $j$ such that $\beta_j=\tau$ then we have: \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s) \land J=\{j\}
\]
\end{enumerate}
\end{definition}
The first statement of the definition states that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a $\tau$ action from a hole. The condition $J=\{j\}$ is a bit restrictive, it could safely be replaced by $\forall j\in J.\, \beta_j=\tau$, allowing the other holes to perform $\tau$ transitions too (because these $\tau$ actions cannot be observed).

@@

\subsubsection*{{\bf Lemma~\ref{lem-decompose}}: Open transition decomposition} ~\\
 Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then, there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation
%\footnote{$\uplus$ is defined on sets as the disjoint union 
% of the contained indexed 
% sets} 
$\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
 t=t'\uplus s'' \land s''\in \mathcal{S}_P \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
   where $\mathcal{S}_P$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{i}^{i \in L_P\uplus L}},\ostate{{t}_{i}^{i \in L'}\uplus 
 	{s}_{i}^{i \in L_P}}|\Pred_{s,t})\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\openrule
 {
 	\beta_j^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{i}^{i \in L_P\uplus L}}\OTarrow {\alpha} \ostate{{s'}_{i}^{~i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\openrule
 	{
 		\beta_j^{j\in J'}, 
 		\Pred\,',  
 		\Post\,'}
 	{\ostate{s_{i}^{i\in L_P}} \OTarrow {\alpha}
 		\ostate{s_{i}'^{\,i\in L_P}}}
 }\\
 Q\models{\openrule
 	{
 		\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post\,''}
 	{\ostate{s_{i}^{i\in L}} \OTarrow {\alpha_Q}
 		\ostate{s_{i}'^{\,i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\iff \Pred\,'
 \land \Pred\,''\land \alpha_Q= \beta_{j_0}$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{i}^{i \in 
 		L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\openrule
 {
 	\beta_{j x}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{t_{i}^{i\in L'}} \OTarrow {\alpha_x} \ostate{t_{i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
 		L'}}|\Pred_{s x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred_{s,t} \land \Pred\,''
 \implies$\\ $\displaystyle{\bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). \beta_j\!=\!\beta_{jx}\!  \land\! 
 \Pred_{OT_x}
 \land \alpha_Q\!=\!\alpha_x \land  
 \Pred_{s, x}\subst{\Post\,''\uplus\Post_{OT_x}}\right)}$

@@

        Let $\Leaves(Q)=p_l^{l\in L_Q}$, 
$\Leaves(P)=p_l^{l\in L}$, $\Leaves(P')={p'}_l^{l\in L'}$.
	Consider $P$ FH-bisimilar to $P'$. It means that there is a relation 
	$\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
	We will consider the relation $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
	t=t'\uplus s'' \land s\in \mathcal{S}_Q \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
	where $\mathcal{S}_Q$ is the set of states of the open automaton of $Q$.	We will prove 
	that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
	states: $(\ostate{s_{1 i}^{i \in L\uplus L_Q}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
	{s}_{1 i}^{i \in L_Q}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
%	automata of 
%	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
Consider an 
	open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
%	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
%	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
\\[-2ex]     
	\[P[Q]_{j_0}\models\openrule
	{
		\beta_j^{j\in J},\Pred_{OT},\Post_{OT}}
	{ \ostate{s_{i}^{i \in L\uplus L_Q}}\OTarrow {\alpha} \ostate{{s'}_{i}^{~i \in 
	L\uplus 
	L_Q}}}\]
Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
	Lemma~\ref{lem-decompose} we have :\\[-2ex]
			\begin{mathpar}
				P\models{\openrule
				{
					\beta_j^{j\in J'}, 
					\Pred\,',  
					\Post\,'}
				{\ostate{s_{1}^{i\in L}} \OTarrow {\alpha}
					\ostate{s_{i}'^{~ i\in L}}}
			}\\
			Q\models{\openrule
				{
					\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
					\Post\,''}
				{\ostate{s_{i}^{i\in L_Q}} \OTarrow {\alpha_Q}
					\ostate{s_{i}'^{~ i\in L_Q}}}
			}\end{mathpar}
			and  $\Pred_{OT} \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post_{OT}=\Post\,'\uplus 
			\Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
			$\Leaves(Q)$). As $P$ is FH-bisimilar to $P'$ and $(\ostate{s_{i}^{i \in 
			L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
			$OT'_x$ 
			of 	open transitions of the automaton of $P'$ such that\\[-2ex] 
			\begin{mathpar}
			%    \left( fresh \ \set{a_i}, \set{b_j}, v_x.\ \
			P'\models\openrule
			{
				\beta_{j x}^{j\in J'}, 
				\Pred_{OT_x},\Post_{OT_x}}
			{\ostate{t_{i}^{i\in L'}} \OTarrow {\alpha_x} \ostate{t_{i x}^{i\in L'}}}
			%         \right)
			\end{mathpar}
			and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
			L'}}|\Pred_{s x})\in 
			\mathcal{R}$; 
			and  \\
			$\displaystyle{\Pred_{s,t} \land \Pred\,'
		 \!\!\implies\!\!\! \bigvee_{x\in X}\!
			\left( \forall j\in J'. \beta_j=\beta_{jx}\!  \land\! 
			\Pred_{OT_x}\!
			\land\! \alpha\!=\!\alpha_x\! \land\!  
			\Pred_{s,x}\subst{\Post\,'\uplus\Post_{OT_x}}\right)}$
			%     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x}) \right)$.

@@

\subsection{Composition properties}\label{sec:app-composition}
This section gives decomposition/composition lemmas and their proofs, these are the equivalent of the composition lemmas for open transitions, but applied to weak open automata.
\begin{lemma}[Weak open transition decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{such that~\qquad}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) 
\text{, }
 \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n (\gamma_{p j}^{j\in J_p})\subst{\bigotimes_{i=p-1}^{0}\Post_i}, \\
{~\hspace{2cm}}\Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} = \beta_p\land \Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} ),
\\
{~\hspace{2cm}}\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p,   \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\
		\end{mathpar} 
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

>>>>

We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:
\begin{lemma}[Open transition decomposition\label{lem-decompose}] Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$Q$.
\end{lemma}

@@

More precisely, we state that $\tau$ is not observable if the automaton always allows any $\tau$ transition from holes, and additionally the global transition resulting from a $\tau$ action of a hole is a $\tau$ transition not changing the pNet's state.
We define $\Id(V)$ as the identity function on variables of $V$.
\begin{definition}[Non-observability of $\tau$ actions]\label{def:Non-ObsTau}
An open automaton $A = <J,\mathcal{S},s_0,V,\mathcal{T}>$ \emph{cannot observe $\tau$ actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have:
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(V)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\beta_j$, $J$,  $\alpha$,  $s$, $s'$, $\Pred$, $\Post$  such that
\[ \openrule
         {
           \beta_j^{j\in J},\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] If there exists $j$ such that $\beta_j=\tau$ then we have: \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(V) \land J=\{j\}
\]
\end{enumerate}
\end{definition}
The first statement of the definition states that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a $\tau$ action from a hole. The condition $J=\{j\}$ is a bit restrictive, it could safely be replaced by $\forall j\in J.\, \beta_j=\tau$, allowing the other holes to perform $\tau$ transitions too (because these $\tau$ actions cannot be observed).

@@

\subsubsection*{{\bf Lemma~\ref{lem-decompose}}: Open transition decomposition} ~\\
 Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then, there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables  
		$\vars(Q)$.

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation
%\footnote{$\uplus$ is defined on sets as the disjoint union 
% of the contained indexed 
% sets} 
$\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
 t=t'\uplus s'' \land s''\in \mathcal{S}_P \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
   where $\mathcal{S}_P$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{i}^{i \in L_P\uplus L}},\ostate{{t}_{i}^{i \in L'}\uplus 
 	{s}_{i}^{i \in L_P}}|\Pred_{s,t})\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\openrule
 {
 	\beta_j^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{i}^{i \in L_P\uplus L}}\OTarrow {\alpha} \ostate{{s'}_{i}^{~i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\openrule
 	{
 		\beta_j^{j\in J'}, 
 		\Pred\,',  
 		\Post\,'}
 	{\ostate{s_{i}^{i\in L_P}} \OTarrow {\alpha}
 		\ostate{s_{i}'^{\,i\in L_P}}}
 }\\
 Q\models{\openrule
 	{
 		\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post\,''}
 	{\ostate{s_{i}^{i\in L}} \OTarrow {\alpha_Q}
 		\ostate{s_{i}'^{\,i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\iff \Pred\,'
 \land \Pred\,''\land \alpha_Q= \beta_{j_0}$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over $\vars(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{i}^{i \in 
 		L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\openrule
 {
 	\beta_{j x}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{t_{i}^{i\in L'}} \OTarrow {\alpha_x} \ostate{t_{i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
 		L'}}|\Pred_{s x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred_{s,t} \land \Pred\,''
 \implies$\\ $\displaystyle{\bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). \beta_j\!=\!\beta_{jx}\!  \land\! 
 \Pred_{OT_x}
 \land \alpha_Q\!=\!\alpha_x \land  
 \Pred_{s, x}\subst{\Post\,''\uplus\Post_{OT_x}}\right)}$

@@

        Let $\Leaves(Q)=p_l^{l\in L_Q}$, 
$\Leaves(P)=p_l^{l\in L}$, $\Leaves(P')={p'}_l^{l\in L'}$.
	Consider $P$ FH-bisimilar to $P'$. It means that there is a relation 
	$\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
	We will consider the relation $\mathcal{R}'=\{(s,t|\Pred_{s,t})|s=s'\uplus s'' \land 
	t=t'\uplus s'' \land s\in \mathcal{S}_Q \land (s',t'|\Pred_{s,t})\in\mathcal{R}\}$ 
	where $\mathcal{S}_Q$ is the set of states of the open automaton of $Q$.	We will prove 
	that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
	states: $(\ostate{s_{1 i}^{i \in L\uplus L_Q}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
	{s}_{1 i}^{i \in L_Q}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
%	automata of 
%	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
Consider an 
	open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
%	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
%	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
\\[-2ex]     
	\[P[Q]_{j_0}\models\openrule
	{
		\beta_j^{j\in J},\Pred_{OT},\Post_{OT}}
	{ \ostate{s_{i}^{i \in L\uplus L_Q}}\OTarrow {\alpha} \ostate{{s'}_{i}^{~i \in 
	L\uplus 
	L_Q}}}\]
Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
	Lemma~\ref{lem-decompose} we have :\\[-2ex]
			\begin{mathpar}
				P\models{\openrule
				{
					\beta_j^{j\in J'}, 
					\Pred\,',  
					\Post\,'}
				{\ostate{s_{1}^{i\in L}} \OTarrow {\alpha}
					\ostate{s_{i}'^{~ i\in L}}}
			}\\
			Q\models{\openrule
				{
					\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
					\Post\,''}
				{\ostate{s_{i}^{i\in L_Q}} \OTarrow {\alpha_Q}
					\ostate{s_{i}'^{~ i\in L_Q}}}
			}\end{mathpar}
			and  $\Pred_{OT} \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post_{OT}=\Post\,'\uplus 
			\Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over $\vars(Q)$). As $P$ is FH-bisimilar to $P'$ and $(\ostate{s_{i}^{i \in 
			L}},\ostate{{t}_{i}^{i \in L'}}|\Pred_{s,t})\in\mathcal{R}$ there is a family 
			$OT'_x$ 
			of 	open transitions of the automaton of $P'$ such that\\[-2ex] 
			\begin{mathpar}
			%    \left( fresh \ \set{a_i}, \set{b_j}, v_x.\ \
			P'\models\openrule
			{
				\beta_{j x}^{j\in J'}, 
				\Pred_{OT_x},\Post_{OT_x}}
			{\ostate{t_{i}^{i\in L'}} \OTarrow {\alpha_x} \ostate{t_{i x}^{i\in L'}}}
			%         \right)
			\end{mathpar}
			and  $\forall x, (\ostate{s_{i}^{i\in L}},\ostate{t_{i x}^{i\in 
			L'}}|\Pred_{s x})\in 
			\mathcal{R}$; 
			and  \\
			$\displaystyle{\Pred_{s,t} \land \Pred\,'
		 \!\!\implies\!\!\! \bigvee_{x\in X}\!
			\left( \forall j\in J'. \beta_j=\beta_{jx}\!  \land\! 
			\Pred_{OT_x}\!
			\land\! \alpha\!=\!\alpha_x\! \land\!  
			\Pred_{s,x}\subst{\Post\,'\uplus\Post_{OT_x}}\right)}$
			%     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x}) \right)$.

@@

\subsection{Composition properties}\label{sec:app-composition}
This section gives decomposition/composition lemmas and their proofs, these are the equivalent of the composition lemmas for open transitions, but applied to weak open automata.
\begin{lemma}[Weak open transition decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{such that~\qquad}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) 
\text{, }
 \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n (\gamma_{p j}^{j\in J_p})\subst{\bigotimes_{i=p-1}^{0}\Post_i}, \\
{~\hspace{2cm}}\Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} = \beta_p\land \Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} ),
\\
{~\hspace{2cm}}\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p,   \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\
		\end{mathpar} 
where for any $p$, $\Post_p$ only acts upon  variables $\vars(Q)$.
\end{lemma}
